\large{\bf Large effects and the infinitesimal model }

### 2023-12-01 review-responses.tex

\begin{point}{}
    I think it is also not correct to claim that ``Since models can be easily set up for which trait distributions are not Gaussian, the implication of this is that for such models, independence of the Mendelian sampling term is not likely a good assumption.'' \revref. As far as I understand, a non-Gaussian distribution of trait values can arise quite easily under the infinitesimal model (where the Mendelian segregation term is independent of the midparent value) e.g, due to strong or non-Gaussian selection or migration or some combination of the two (see, e.g., Figure 1 in Barton et al).
\end{point}

\reply{
    Thanks -- we've changed this to ``since models can be easily set up for which the distribution
    of Mendelian sampling terms is not Gaussian, \ldots''
    and provided additional context \revref.
}

\begin{point}{}
    Section 4, first paragraph \revref: I don't think the third assumption (that trait values in the population follow a Gaussian distribution) is a component of the ``infinitesimal model''. I think Turelli (2017) is also quite clear about this stating that a Gaussian distribution of trait values should only emerge for Gaussian or very weak selection on the population.
\end{point}

\reply{This  is almost verbatim from the abstract of Turelli 2017:
\begin{quote}
This commentary distinguishes three nested approximations, referred to as “infinitesimal genetics,” “Gaussian descendants” and “Gaussian population,” each plausibly called “the infinitesimal model.” 
\end{quote} 
}

@@

\reply{
    Yes; we've added this to Figure~\ref{fig:exponent_hist}.
}

\begin{point}{\revref}
    Sentence starting ``To see what knowing the parental trait value...'': In what sense is this the ``parental'' trait value??
\end{point}

\reply{
}

\begin{point}{\revref}
    Last inequality on Page 13: should the subscript be $j$ instead of $i$?
\end{point}

\reply{It should be $\ell$, and is now corrected \revref.  Thanks for noticing the discrepancy; is an artefact of an earlier version which used $i$ for an index on loci.
}

@@

\reply{
    Good points; both are fixed: \revref and \llname{loci_polymorphic}.
}

\begin{point}{}
p 7, Sect 4: As noted above, please describe this model in more detail to distinguish it
from that in Barton et al (2017), which makes much weaker assumptions. Note that it is
not justifiable except under quite extreme assumptions.
\end{point}

\reply{
}

\begin{point}{\revref}
Third display equation in Sect 4.1 \revref: The index k in the product should start at k = 0.
\end{point}

\reply{Thank you for noticing this.  We have corrected this, and the subsequent indices \revref.
}

\begin{point}{\revref}
p 8, last line: Delete ‘to’
\end{point}

\reply{Done \revref.
}

### 2023-12-01 writeup.tex
minor add
More minor edits.

The Gaussian distribution appears thanks to ``the'' Central Limit Theorem,
i.e., the universality of adding up a great many very small and independent things.
However, there is a wider class of Central Limit Theorems,
which describe the distributions of sums of a great many independent and exchangeable things:
the \emph{stable laws} \citep{gnedenko1968limit}.
Roughly speaking, these say that if we add together $n$ independent copies of a random variable $X$
for which $\P\{ X > t \}$ is proportional to $t^{-\alpha}$ for some $0 < \alpha < 2$,
then -- almost regardless of the details -- the result, scaled by $n^{1/\alpha}$,
is well-approximated by a universal form, an $\alpha$-stable distribution
(and the approximation becomes exact, as $n \to \infty$).
Oth the other hand, if $\alpha \ge 2$ then the distribution has finite variance,
and so the sum of $n$, divided by $\sqrt{n}$, converges to a Gaussian distribution.
In such ensembles, single large entries can be important:
concretely, the largest of the $n$ copies will be of order $n^{1/\alpha}$,
so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.
On the other hand, the largest of $n$ Gaussians is of order $\sqrt{\log n}$. \revpoint{2}{1}

@@

First, consider the case without selection, so $\mu$ is a constant
which we may take equal to 1.  Then, our condition on the characteristic function simplifies to
\begin{equation}\label{eq:CF_neutral}
	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta) %e^{-\frac{\eta^{2}\theta^{2}}{2}} 
	= \phi(\theta) ,
\end{equation}
\ie $\phi(\theta)$ is a fixed point of $F$, where
 \[
	F(\phi)(\theta) = {\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta),
\]
whereas any fixed point must be the limit of an arbitrary characteristic function, $\varphi(\theta)$, under repeated applications of $F$,
\ie \revpoint{2}{7}
\begin{align*}
	\phi(\theta) 
    &= \lim_{n \to \infty} F^{(n)}(\varphi)(\theta) \\
    &= \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=1}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}} .
\end{align*}
% Iterating $F$ $n$ times, we get
% \[
% 	F^{(n)}(\varphi)(\theta) =  {\textstyle \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}}\prod_{k=1}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}},
% \]
% so that
% \[
% 	\phi(\theta) = \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=1}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}
% \]
is a fixed point whenever this limit exists.

Since $\varphi$ and $\Phi$ are characteristic functions, they are continuous and $\phi(0) = \Phi(0) = 1$, and thus both are non-zero in a neighbourhood of 0. In particular, $\psi(\theta) = \Log{\varphi}(\theta)$ and $\Psi(\theta) = \Log{\Phi(\theta)}$ exist in some neighbourhood of 0 as well\footnote{We use $\Log$ to denote the principal value of the complex logarithm, for which the imaginary part always lies in the interval $(-\pi,\pi]$.}.  From the above, we have
\[
	\Log{F^{(n)}(\varphi)(\theta)} = {\textstyle 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)} 
		+ \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}.
\]
We briefly explore what possible limiting values are valid characteristic functions.  We refer to the first term on the left hand side as the \emph{reproduction term} and the remaining sum as the \emph{noise term}, and only consider to the case when both have well-posed limits independent of the other. \llabel{limit_remark}

\paragraph{Noise Terms}
First, consider the noise terms.  A necessary condition for $\sum_{k=1}^{\infty} 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)$ to converge is that $2^{k} \Psi\left(\frac{\theta}{2^{k}}\right) \to 0$; since $\Psi(0) = 0$, this implies that if it exists, $\Psi'(0+) = 0$.  In particular, as we would expect, if $R$ has a mean, it must be 0.  If, on the other hand, $\Psi(\theta)$ is twice differentiable, so $R$ has mean 0 and finite variance, say $\Psi''(0) = \eta^{2}$, then Taylor's theorem tells us that for $\theta$ sufficiently small, 
\[
	\Psi(\theta) = \frac{\eta^{2}}{2} \theta^{2} + r(\theta)\theta^{2},
\]
where $r(\theta) \to 0$ as $\theta \to 0$.  Thus,
\begin{equation*}
	\frac{\textstyle 2^{k+1} \Psi\left(\frac{\theta}{2^{k+1}}\right)}{\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	= \frac{1}{2} \frac{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k+1}}\right)}\theta^{2}}{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k}}\right)}\theta^{2}}
	\to \frac{1}{2},
\end{equation*}
as $k \to \infty$ and the sum converges by the ratio test.  In this case,
\begin{align*}
	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	 &=  \sum_{k=1}^{n-1} 2^{k} {\textstyle \left(- \frac{\eta^{2}}{2}  \frac{\theta^{2}}{2^{2k}}
		+ r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{2k}}\right)}\\
	&=- \eta^{2} \theta^{2} (1 - 2^{-n+1})  + \sum_{k=1}^{n-1} {\textstyle r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{k}}}
\end{align*}
If we assume that $r(\theta) \equiv 0$, so that the noise is $N(0,\eta^{2})$, then the limiting distribution is $N(0,2\eta^{2})$. When  $r(\theta)$ is non-zero, we don't expect the higher order terms to vanish.

It is unclear what more can be said in general. If however, we specialize to the case when $R$ is a stable law, we can explicitly compute the limiting noise term.  We recall that if $\Psi(\theta)$ corresponds to a stable law (see \eg \cite{breiman1992probability}), then
\[
	\Psi(\theta) = im\theta - c|\theta|^{\alpha} (1-i\beta \sgn(\theta)\omega_{\alpha}(\theta))
\]
for $\alpha \in (0,2]$, $\beta \in [-1,1]$, $c \geq 0$, and $m \in \mathbb{R}$ and
\[
	\omega_{\alpha}(\theta) = \begin{cases}
		\tan{\frac{\pi\alpha}{2}} & \text{if $\alpha \neq 1$, and}\\
		-\frac{2}{\pi}\log{|\theta|} & \text{if $\alpha = 1$.}
	\end{cases}
\]
By our general remarks above, we can assume $m = 0$ and exclude the case $\alpha = 1$, $\beta = 0$, for which  $\Psi'(0+) = c$. We then have
\begin{align*}
	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	 &=  \sum_{k=1}^{n-1} -2^{(1-\alpha)k} c|\theta|^{\alpha} {\textstyle\left(1-i \beta\sgn\left(\frac{\theta}{2^{k}}\right)
	 	 \omega_{\alpha}\left(\frac{\theta}{2^{k}}\right)\right)}\\
	&= \begin{cases}
		 -c|\theta|^{\alpha}\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}\left(1-i\beta \sgn(\theta) \tan{\frac{\pi\alpha}{2}}\right) & \text{if $\alpha \neq 1$, and}\\	
         -(n-1) c|\theta| \left(1+\frac{2i\beta}{\pi}\sgn(\theta)\left(\log{|\theta|} - \frac{n}{2} \log{2}\right)\right) & \text{if $\alpha = 1$,}
	\end{cases}
\end{align*}
which converges as $n \to \infty$ if and only if $\alpha > 1$, in which case the limit is 
\begin{equation}\label{eq:stablenoises}
	 -\frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}\left(1-i\beta \sgn(\theta) \tan{\frac{\pi\alpha}{2}}\right), 
\end{equation}
which corresponds to a $\left(\alpha,\beta, \frac{c}{1-2^{1-\alpha}},0\right)$-stable law.

@@

Since $\varphi(\theta)$ is a characteristic function, it has the Hermitian property: $\phi(-\theta) = \overline{\phi(\theta)}$, and thus $\psi(-\theta) = \overline{\psi(\theta)}$ and $\alpha(-\theta) = \overline{\alpha(\theta)}$.  In particular, if one (and by extension both) limits exist, we must have $\alpha'(0-) = \overline{\alpha'(0+)}$.  In this case, setting $-\gamma + i m = \alpha'(0+)$, we have
\[
	\alpha(\theta) = \begin{cases} 
		i m\theta - \gamma \theta & \text{if $\theta \geq 0$, and}\\
		i m\theta + \gamma \theta & \text{if $\theta < 0$}.
	\end{cases}
\]
If $\gamma > 0$, then we recognize $e^{\alpha(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median 0 and median absolute deviation $\gamma$, whereas if $\gamma = 0$, then $e^{\alpha(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.
(And, $\gamma < 0$ does not give a valid characteristic function.)
% % I think we've shown this characteristic function has to be exp(im\theta - \gamma|\theta|);
% % if $\gamma < 0$ then it is not a characteristic function; am I missing something?
% In general, it does not seem possible to exclude $\gamma < 0$. However, recalling that if $\phi(\theta)$ is the characteristic function of a random variable $Z$, Jensen's inequality tells us that for all $\theta$,
% \begin{equation}\label{eq:CF-J}
% 	|\phi(\theta)| = \left|\E\left[e^{i\theta Z}\right]\right| \leq \E\left[\left|e^{i\theta Z}\right|\right] = 1,
% \end{equation}
% we can conclude $\gamma \geq 0$ when the limiting noise term is a mixture of Gaussians and stable laws of the form \eqref{eq:stablenoises}: we then have 
% \[
% 	|\phi(\theta)| = e^{\re(\alpha(\theta) + o(|\theta|))} = e^{\gamma |\theta| + o(|\theta|))},
% \] 
% and thus is less than or equal to one in a small neighbourhood of $0$ if and only if $\gamma < 0$.

@@

If, however, we consider not any one locus, but rather all loci, we get a different picture. 
The variables $\eta^{[j]}_{\ell}$ are independent, and thus so are the Bernoulli random variables 
\begin{equation}
	\1^{M}_{\ell} = \begin{cases}
        1 & \text{if $\frac{\eta^{[j]}_{\ell}}{\bar{\gamma} M} > \epsilon$, and}\\
		0 & \text{otherwise.}
	\end{cases}
\end{equation}
Now, since
\[
	\lim_{M \to \infty} \sum_{\ell = 1}^{M} \frac{2\gamma_{\ell}}{\pi \bar{\gamma} M \epsilon} = \frac{2}{\pi \epsilon},
\]
the number of loci at which the magnitude of the effect size exceeds $M \epsilon$, which is 
$\sum_{\ell = 1}^{M} \1^{M}_{\ell}$,
converges weakly as $M \to \infty$ to a Poisson random variable with mean $2/(\pi \epsilon)$ (see \eg Theorem 2.6.1 in \cite{Durrett2005}).  In particular, for large values of $M$ and fixed $\epsilon > 0$,
\begin{align*}
    \P\left\{\max_{i=1}^{M} \frac{\eta^{[j]}_{\ell}}{\bar{\gamma} M} > \epsilon\right\}  \sim 1-e^{-\frac{2}{\pi \bar{\gamma} \epsilon}},
\end{align*}
which approaches 1 as $\epsilon \to 0$, while the expected number of $M$ alleles such that $\frac{\eta^{[j]}_{\ell}}{\bar{\gamma} M} > \epsilon$ is asymptotically $\frac{2}{\pi\epsilon}$.

>>>>

The Gaussian distribution appears thanks to ``the'' Central Limit Theorem,
i.e., the universality of adding up a great many very small and independent things.
However, there is a wider class of Central Limit Theorems,
which describe the distributions of sums of a great many independent and exchangeable things:
the \emph{stable laws} \citep{gnedenko1968limit}.
Roughly speaking, these say that if we add together $n$ independent copies of a random variable $X$
for which $\P\{ X > t \}$ is proportional to $t^{-\alpha}$ for some $0 < \alpha < 2$,
then -- almost regardless of the details -- the result, scaled by $n^{1/\alpha}$,
is well-approximated by a universal form, an $\alpha$-stable distribution
(and the approximation becomes exact, as $n \to \infty$).
Oth the other hand, if $\alpha \ge 2$ then the distribution has finite variance,
and so the sum of $n$, divided by $\sqrt{n}$, converges to a Gaussian distribution.
In such ensembles, single large entries can be important:
concretely, the largest of the $n$ copies will be of order $n^{1/\alpha}$,
so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.
On the other hand, the largest of $n$ Gaussians is of order $\sqrt{\log n}$. \revpoint{2}{1} \revpoint{1}{11}

@@

First, consider the case without selection, so $\mu$ is a constant
which we may take equal to 1.  Then, our condition on the characteristic function simplifies to
\begin{equation}\label{eq:CF_neutral}
	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta) %e^{-\frac{\eta^{2}\theta^{2}}{2}} 
	= \phi(\theta) ,
\end{equation}
\ie $\phi(\theta)$ is a fixed point of $F$, where
 \[
	F(\phi)(\theta) = {\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta),
\]
whereas any fixed point must be the limit of an arbitrary characteristic function, $\varphi(\theta)$, under repeated applications of $F$,
\ie \revpoint{2}{7}
\begin{align*}
	\phi(\theta) 
    &= \lim_{n \to \infty} F^{(n)}(\varphi)(\theta) \\
    &= \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}} .
\end{align*}
% Iterating $F$ $n$ times, we get
% \[
% 	F^{(n)}(\varphi)(\theta) =  {\textstyle \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}}\prod_{k=0}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}},
% \]
% so that
% \[
% 	\phi(\theta) = \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}
% \]
is a fixed point whenever this limit exists.

Since $\varphi$ and $\Phi$ are characteristic functions, they are continuous and $\phi(0) = \Phi(0) = 1$, and thus both are non-zero in a neighbourhood of 0. In particular, $\psi(\theta) = \Log{\varphi}(\theta)$ and $\Psi(\theta) = \Log{\Phi(\theta)}$ exist in some neighbourhood of 0 as well\footnote{We use $\Log$ to denote the principal value of the complex logarithm, for which the imaginary part always lies in the interval $(-\pi,\pi]$.}.  From the above, we have
\[
	\Log{F^{(n)}(\varphi)(\theta)} = {\textstyle 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)} 
		+ \sum_{k=0}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}.
\]
We briefly explore what possible limiting values are valid characteristic functions.  We refer to the first term on the left hand side as the \emph{reproduction term} and the remaining sum as the \emph{noise term}, and only consider the case \revpoint{2}{8} when both have well-posed limits independent of the other. \llabel{limit_remark} % We don't need to limit to this caee.

\paragraph{Noise Terms}
First, consider the noise terms.  A necessary condition for $\sum_{k=0}^{\infty} 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)$ to converge is that $2^{k} \Psi\left(\frac{\theta}{2^{k}}\right) \to 0$; since $\Psi(0) = 0$, this implies that if it exists, $\Psi'(0+) = 0$.  In particular, as we would expect, if $R$ has a mean, it must be 0.  If, on the other hand, $\Psi(\theta)$ is twice differentiable, so $R$ has mean 0 and finite variance, say $\Psi''(0) = \eta^{2}$, then Taylor's theorem tells us that for $\theta$ sufficiently small, 
\[
	\Psi(\theta) = \frac{\eta^{2}}{2} \theta^{2} + r(\theta)\theta^{2},
\]
where $r(\theta) \to 0$ as $\theta \to 0$.  Thus,
\begin{equation*}
	\frac{\textstyle 2^{k+1} \Psi\left(\frac{\theta}{2^{k+1}}\right)}{\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	= \frac{1}{2} \frac{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k+1}}\right)}\theta^{2}}{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k}}\right)}\theta^{2}}
	\to \frac{1}{2},
\end{equation*}
as $k \to \infty$ and the sum converges by the ratio test.  In this case,
\begin{align*}
	 \sum_{k=0}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	 &=  \sum_{k=0}^{n-1} 2^{k} {\textstyle \left(- \frac{\eta^{2}}{2}  \frac{\theta^{2}}{2^{2k}}
		+ r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{2k}}\right)}\\
	&=- \eta^{2} \theta^{2} (1 - 2^{-n+1})  + \sum_{k=0}^{n-1} {\textstyle r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{k}}}
\end{align*}
If we assume that $r(\theta) \equiv 0$, so that the noise is $N(0,\eta^{2})$, then the limiting distribution is $N(0,2\eta^{2})$. When  $r(\theta)$ is non-zero, we don't expect the higher order terms to vanish.

It is unclear what more can be said in general. If however, we specialize to the case when $R$ is a stable law, we can explicitly compute the limiting noise term.  We recall that if $\Psi(\theta)$ corresponds to a stable law (see \eg \cite{breiman1992probability}), then
\[
	\Psi(\theta) = im\theta - c|\theta|^{\alpha} (1-i\beta \sgn(\theta)\omega_{\alpha}(\theta))
\]
for $\alpha \in (0,2]$, $\beta \in [-1,1]$, $c \geq 0$, and $m \in \mathbb{R}$ and
\[
	\omega_{\alpha}(\theta) = \begin{cases}
		\tan{\frac{\pi\alpha}{2}} & \text{if $\alpha \neq 1$, and}\\
		-\frac{2}{\pi}\log{|\theta|} & \text{if $\alpha = 1$.}
	\end{cases}
\]
By our general remarks above, we can assume $m = 0$ and exclude the case $\alpha = 1$, $\beta = 0$, for which  $\Psi'(0+) = c$. We then have
\begin{align*}
	 \sum_{k=0}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	 &=  \sum_{k=0}^{n-1} -2^{(1-\alpha)k} c|\theta|^{\alpha} {\textstyle\left(1-i \beta\sgn\left(\frac{\theta}{2^{k}}\right)
	 	 \omega_{\alpha}\left(\frac{\theta}{2^{k}}\right)\right)}\\
	&= \begin{cases}
		 -c|\theta|^{\alpha}\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}\left(1-i\beta \sgn(\theta) \tan{\frac{\pi\alpha}{2}}\right) & \text{if $\alpha \neq 1$, and}\\	
         -(n-1) c|\theta| \left(1+\frac{2i\beta}{\pi}\sgn(\theta)\left(\log{|\theta|} - \frac{n}{2} \log{2}\right)\right) & \text{if $\alpha = 1$,}
	\end{cases}
\end{align*}
which converges as $n \to \infty$ if and only if $\alpha > 1$, in which case the limit is 
\begin{equation}\label{eq:stablenoises}
	 -\frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}\left(1-i\beta \sgn(\theta) \tan{\frac{\pi\alpha}{2}}\right), 
\end{equation}
which corresponds to a $\left(\alpha,\beta, \frac{c}{1-2^{1-\alpha}},0\right)$-stable law.

@@

Since $\varphi(\theta)$ is a characteristic function, it has the Hermitian property: $\phi(-\theta) = \overline{\phi(\theta)}$, and thus $\psi(-\theta) = \overline{\psi(\theta)}$ and $\alpha(-\theta) = \overline{\alpha(\theta)}$.  In particular, if one (and by extension both) limits exist, we must have $\alpha'(0-) = \overline{\alpha'(0+)}$.  In this case, setting $-\gamma + i m = \alpha'(0+)$, we have
\[
	\alpha(\theta) = \begin{cases} 
		i m\theta - \gamma \theta & \text{if $\theta \geq 0$, and}\\
		i m\theta + \gamma \theta & \text{if $\theta < 0$}.
	\end{cases}
\]
If $\gamma > 0$, then we recognize $e^{\psi(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median 0 and median absolute deviation $\gamma$, whereas if $\gamma = 0$, then $e^{\psi(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.

% This might seem a bit roundabout, but the point is that we can prove that $e^{\psi(\theta)}$ is -- in this case -- a characteristic function, so we don't need to lose generality by assuming that it is one.  Our goal is to ask when $\lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}$ is a valid characteristic function, and to see what it can be.  
To exclude the case $\gamma < 0$, we first observe that if, as we assume, $\varphi(\theta)$ is the characteristic function of some random variable, say $X$, then $\varphi_{n}(\theta) = \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}} = e^{2^{n}\psi\left(\frac{\theta}{2^{n}}\right)}$ is the characteristic function of $2^{-n} \sum_{i = 1}^{2^{n}} X_{i}$, where $X_{1},\ldots,X_{2^{n}}$ are i.i.d. copies of $X$; now, L\'evy's continuity theorem (see \eg Theorem 2.3.4 in \cite{Durrett2005}) tells us that if (i) $\phi_{n}(\theta)$ has a point-wise limit, say $\phi_{\infty}(\theta)$, and (ii) $\phi_{\infty}(\theta)$ is continuous at 0, then $\phi_{\infty}(\theta)$ is the characteristic function of \emph{some} random variable; in particular, if $\alpha(\theta)$ exists and is bounded (\eg if the left and right-hand side derivatives exist, as above), then $\psi(\theta) = \alpha(\theta)\theta$ is continuous at 0, and hence so is $e^{\psi(\theta)}$.  Thus, in this case $e^{\psi(\theta)}$ is itself a characteristic function.  Moreover, being a characteristic function, $e^{\psi(\theta)}$ must satisfy $|e^{\psi(\theta)}| = e^{\Re \psi(\theta)} \leq 1$, whence $\gamma \geq 0$.

% % I think we've shown this characteristic function has to be exp(im\theta - \gamma|\theta|);
% % if $\gamma < 0$ then it is not a characteristic function; am I missing something?
% In general, it does not seem possible to exclude $\gamma < 0$. However, recalling that if $\phi(\theta)$ is the characteristic function of a random variable $Z$, Jensen's inequality tells us that for all $\theta$,
% \begin{equation}\label{eq:CF-J}
% 	|\phi(\theta)| = \left|\E\left[e^{i\theta Z}\right]\right| \leq \E\left[\left|e^{i\theta Z}\right|\right] = 1,
% \end{equation}
% we can conclude $\gamma \geq 0$ when the limiting noise term is a mixture of Gaussians and stable laws of the form \eqref{eq:stablenoises}: we then have 
% \[
% 	|\phi(\theta)| = e^{\re(\alpha(\theta) + o(|\theta|))} = e^{\gamma |\theta| + o(|\theta|))},
% \] 
% and thus is less than or equal to one in a small neighbourhood of $0$ if and only if $\gamma < 0$.

@@

If, however, we consider not any one locus, but rather all loci, we get a different picture. 
The variables $\eta^{[j]}_{\ell}$ are independent, and thus so are the Bernoulli random variables 
\begin{equation}
	\1^{M}_{\ell} = \begin{cases}
        1 & \text{if $\frac{\eta^{[j]}_{\ell}}{\bar{\gamma} M} > \epsilon$, and}\\
		0 & \text{otherwise.}
	\end{cases}
\end{equation}
Now, since
\[
	\lim_{M \to \infty} \sum_{\ell = 1}^{M} \frac{2\gamma_{\ell}}{\pi \bar{\gamma} M \epsilon} = \frac{2}{\pi \epsilon},
\]
the number of loci at which the magnitude of the effect size exceeds $M \epsilon$, which is 
$\sum_{\ell = 1}^{M} \1^{M}_{\ell}$,
converges weakly as $M \to \infty$ to a Poisson random variable with mean $2/(\pi \epsilon)$ (see \eg Theorem 2.6.1 in \cite{Durrett2005}).  In particular, for large values of $M$ and fixed $\epsilon > 0$,\revpoint{1}{17}
\begin{align*}
    \P\left\{\max_{\ell=1}^{M} \frac{\eta^{[j]}_{\ell}}{\bar{\gamma} M} > \epsilon\right\}  \sim 1-e^{-\frac{2}{\pi \bar{\gamma} \epsilon}},
\end{align*}
which approaches 1 as $\epsilon \to 0$, while the expected number of $M$ alleles such that $\frac{\eta^{[j]}_{\ell}}{\bar{\gamma} M} > \epsilon$ is asymptotically $\frac{2}{\pi\epsilon}$.
